---
title: "Assessment 1.1"
author: "Christopher Schmidt S4532362"
date: "15/08/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Markdown editing

This is an **R Markdown** document. *Markdown* is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

[Markdown Basics](https://rmarkdown.rstudio.com/lesson-8.html)

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
library(tidyverse)
library(janitor)
library(skimr)

fastfood <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv')

fastfood%>%head(10) #Display the first 10 rows of the dataset.

knitr::kable(head(fastfood[, 1:5]))

fastfood%>%
  filter(calories>1130) #Display the observations that has more than 1000 calories
```


