---
title: "Assessment 1.1"
author: "Christopher Schmidt S4532362"
date: "15/08/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RMarkdown editing

**R Markdown** contains a combination of markdown (*content with simple text formatting*) and R code chunks.

R Markdown documents allow you to weave together narrative text and code to produce elegantly formatted output in dozens of formats, such as HTML, PDF, and MS Word documents. 

For more details on how to use R Markdown see [Markdown Basics](https://rmarkdown.rstudio.com/lesson-8.html)

This document was created with **R Markdown**.

```{r}
## install packages once via console using - install.packages("tidyverse")
library(tidyverse) # load tidyverse library
library(skimr)     # load skimr library
library(janitor)   # load janitor library
library(ggplot2)   # load ggplot library
```

### Insert an R chunk and create a dataset variable called `fastfood`
```{r}
fastfood <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv')
```

### Display the **first 10 rows** and **first 5 variables** of the `fastfood` dataset
```{r}
# fastfood%>%head(10) # Display the first 10 rows of the dataset using head() function.
knitr::kable(head(fastfood[, 1:5])) # first 10 rows of the dataset and the first 5 variables using kabble() function
```

### Display observations that have more than **1000 calories**
```{r}
fastfood%>%             # Pipe dataset into filter function 
  filter(calories>1000) # Filter by calories greater than 1000
```

### Save observations with more than **40 in total_fat** and more than **80 in total_carb** in a new variable called `dont_eat_this`
```{r}
dont_eat_this <- fastfood%>%           # Create new variable called `dont_eat_this` and assign filtered dataset to it 
  filter(total_fat>40 & total_carb>80) # filter by total_fat greater than 40 & total_carb greatre than 80
```

### Using the initial dataset variable, display observations from restaurant **Mcdonalds**
```{r}
fastfood%>%                       # Pipe dataset into filter function
  filter(restaurant=="Mcdonalds") # Filter by restaurant named Mcdonalds
```

## Part 2

### Display the types of variables in the dataset using `skimr` package
```{r}
#skim(fastfood)
fastfood%>%skim() # Pipe dataset into skim function
```

### Show the count of observations for each restaurant.
```{r}
fastfood%>%count(restaurant) # Pipe dataset into count function and count restaurants.
```

### Using grouppings (group_by()), summarise and display the average number of calories for each restaurant.
```{r}
fastfood%>%
  group_by(restaurant)%>%
  summarise(calories=mean(calories))
```

### Add a variable to the dataset, which calculates the average calories per type of restaurant and call it `average_calories`
```{r}
fastfood<-fastfood %>%
  group_by(restaurant)%>%
  mutate(average_calories=mean(calories))
```

### Display the data vis of total fat per each type of restaurant
```{r}
fastfood%>%
  group_by(restaurant)%>%
  summarise(restaurant_fat=sum(total_fat))%>%
  ggplot(aes(reorder(restaurant, -restaurant_fat),restaurant_fat,color=restaurant,fill=restaurant))+
  geom_col()+
  labs(x = "Restaurant", y = "Total fat per each type of restaurant")+
  coord_flip()
```


### Add a variable to the dataset, which calculates the maximum total fat per type of restaurant and call it `max_total_fat
```{r}
fastfood<-fastfood %>%
  group_by(restaurant)%>%
  mutate(max_total_fat=max(total_fat))
```

### Use observations for Mcdonalds to plot sugar variable against protein with `geom_point()`
```{r}
fastfood%>%
  filter(restaurant=="Mcdonalds") %>%
  ggplot(aes(sugar,protein))+
  geom_point()
```






