---
title: "Assessment 1"
author: "Christopher Schmidt S4532362"
date: "27/08/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RMarkdown editing

**R Markdown** contains a combination of markdown (*content with simple text formatting*) and R code chunks.

R Markdown documents allow you to weave together narrative text and code to produce elegantly formatted output in dozens of formats, such as HTML, PDF, and MS Word documents. 

For more details on how to use R Markdown see [Markdown Basics](https://rmarkdown.rstudio.com/lesson-8.html)

This document was created with **R Markdown**.

```{r}
## install packages once via console using - install.packages("tidyverse")
library(tidyverse) # load tidyverse library
library(skimr)     # load skimr library
library(janitor)   # load janitor library
library(ggplot2)   # load ggplot library
library(knitr)     # load knitr library
```

### 6. Insert an R chunk and create a dataset variable called `fastfood`
```{r}
fastfood <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv')
```

### 7. Display the **first 10 rows** and **first 5 variables** of the `fastfood` dataset
```{r}
fastfood%>%head(10)                 # Display the first 10 rows of the dataset using head() function.

kable((fastfood[1:10, 1:5])) # first 10 rows of the dataset and the first 5 variables using kabble() function

```

### 8. Display observations that have more than **1000 calories**
```{r}
fastfood%>%             # Pipe dataset into filter function 
  filter(calories>1000) # Filter by calories greater than 1000 

```

### 9. Save observations with more than **40 in total_fat** and more than **80 in total_carb** in a new variable called `dont_eat_this`
```{r}
dont_eat_this <- fastfood%>%           # Create new variable called `dont_eat_this` and assign filtered dataset to it 
  filter(total_fat>40 & total_carb>80) # filter by total_fat greater than 40 & total_carb greatre than 80
```

### 10. Using the initial dataset variable, display observations from restaurant **Mcdonalds**
```{r}
fastfood%>%                       # Pipe dataset into filter function
  filter(restaurant=="Mcdonalds") # Filter by restaurant named Mcdonalds
```

## Part 2

### 11. Display the types of variables in the dataset using `skimr` package
```{r}
#skim(fastfood)
fastfood %>% 
  skim() # Pipe dataset into skim function
```

### 12. Show the `count` of observations for each restaurant.
```{r}
fastfood %>% 
  count(restaurant) # Pipe dataset into count function and count restaurants.
```

### 13. Using grouppings `group_by()`, summarise and display the average number of calories for each restaurant.
```{r}
fastfood %>%
  group_by(restaurant) %>% # Pipe dataset into group_by function and group by restaurant
  summarise(calories=mean(calories)) # Pipe dataset into summarise function and use mean function to get avg calories
```

### 14. Add a variable to the dataset, which calculates the average calories per type of restaurant and call it `average_calories`
```{r}
fastfood <- fastfood %>%
  group_by(restaurant) %>% # Pipe dataset into group_by function and group by restaurant
  mutate(average_calories=mean(calories)) #Pipe into mutate function to create new variable, use mean function to get avg calories
```

### 15. Display the data vis of total fat per each type of restaurant

#### Observations
gemo_col was used as we have one discrete and one continuous 
Chick Fil-A has the lowest total fat 
Burger King has the highest total fat

```{r}
fastfood%>%
  group_by(restaurant)%>% # Pipe dataset into group_by function and group by restaurant
  summarise(restaurant_fat=sum(total_fat))%>% # Pipe dataset into summarise function and use sum function to get total fat
  ggplot(aes(reorder(restaurant, -restaurant_fat),restaurant_fat,color=restaurant,fill=restaurant))+ #use ggplot and reorder by fat
  geom_col()+ # add Bar charts using geom_col
  labs(x = "Restaurant", y = "Total fat per restaurant")+ #add label to the plot
  coord_flip()+ # change postion of X and Y coords
  theme(legend.position = "none") # remove the legend
```


### 16. Add a variable to the dataset, which calculates the maximum total fat per type of restaurant and call it `max_total_fat`
```{r}
fastfood<-fastfood %>%  # write back to dataset 
  group_by(restaurant)%>% # Pipe dataset into group_by function and group by restaurant 
  mutate(max_total_fat=max(total_fat)) #Pipe into mutate function to create new variable, use max fuction to get max fat
```

### 17. Use observations for Mcdonalds to plot sugar variable against protein with `geom_point()`
```{r}
fastfood%>%
  filter(restaurant=="Mcdonalds") %>% # Pipe dataset into filter function and select Mcdonalds
  ggplot(aes(sugar,protein))+ # use ggplot to plot sugar variable against protein
  geom_point() # Pipe into geom_point to create scatterplots
```

### PART 3 

### 18. Identify variable(s) which should be factors and transform their type into a factor variable.
```{r}
names(fastfood) #investigate dataset using names function

fastfood <- fastfood %>% # Pipe dataset into mutate  
  mutate(restaurant=as_factor(restaurant)) #convert restaurant to a factor

fastfood %>% head() #check to see if restaurant has been converted to a factor
```

### 19. Create two new variables: add a new variable to the dataset `calories_type`
### Use `calories` variable for `cut_number()` function to split it into 3 categories `n=3`, add labels `labels=c("low", "med", "high")` and make the dataset ordered. 
```{r}
fastfood <- fastfood %>% # Pipe dataset into mutate 
  mutate(calories_type = cut_number(calories, n=3, labels=c("low", "med", "high"))) %>% #use cut_ restaurant to a factor
  arrange(calories) #make dataset ordered using arrange by calories
```

### 20. Display the dataviz of shows the distribution of `calories_type` in food items for each type of restaurant.
### Use facets, coordinates and theme layers to make your data viz visually appealing and meaningful. 
### Use factors related data viz functions
```{r}
fastfood %>% # Pipe dataset into ggplot
  ggplot(aes(calories_type,color=calories_type,fill = calories_type))+ # use ggplot to plot calories_type by restaurant
  geom_bar()+ # add Bar charts using geom_bar
  facet_wrap(vars(restaurant), ncol = 4)+ # use function facet_wrap to create a plot for each restaurant
  labs(x = "Distribution of calories (low, med, high) per Restaurant") #add label to the plot
```

### 21. Add the variable to the dataset and call it `trans_fat_percent`
### `trans_fat_percent` equals the percentage of `trans_fat` in `total_fat`
```{r}
fastfood <- fastfood %>%
  mutate(trans_fat_percent = trans_fat/total_fat*100) # Pipe into mutate to create new variable, use trans_fat/total_fat*100
```

### 22. Display the dataviz of shows the distribution of `trans_fat` in food items for each type of restaurant.
```{r}
fastfood %>% 
  ggplot(aes(trans_fat))+  # use ggplot to plot trans_fat
  geom_histogram(binwidth = 0.3)+ # add histogram using geom_histogram
  facet_wrap(vars(restaurant), ncol = 4) # use function facet_wrap to create a plot for each restaurant

```

### 23. Calculate and show the average (mean) `total_fat` for each type of restaurant.
```{r}
fastfood %>%
  group_by(restaurant)%>% # Pipe dataset into group_by function and group by restaurant
  summarise(total_fat=mean(total_fat)) # Pipe into summarise function and use mean function to show average total fat 
```

### 24. Create a dataviz that allow to compare different restaurants on this variable `total_fat`.
```{r}
fastfood %>%
  group_by(restaurant)%>% # Pipe dataset into group_by function and group by restaurant
  summarise(restaurant_fat=mean(total_fat))%>% # Pipe into summarise function and use mean function to show average total fat 
  ggplot(aes(reorder(restaurant, -restaurant_fat),restaurant_fat, fill=restaurant))+ # use ggplot to plot fat by restaurant
  geom_col()+ # add Bar charts using geom_col
  labs(x = "Restaurant", y = "Average fat per restaurant")+ # #add label to the plot
  theme(legend.position = "none") # remove the legend
```














