---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## *Christopher Schmidt* 
## **S4532362**
******
# Assessment 2 
******

```{r}
## install packages once via console using - install.packages("tidyverse")
library(tidyverse) # load tidyverse library
library(skimr)     # load skimr library
library(janitor)   # load janitor library
library(ggplot2)   # load ggplot library
library(knitr)     # load knitr library

```

### Insert an R chunk and create a dataset variable called `data`
```{r}
data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-11-20/thanksgiving_meals.csv')
```

### Part 2: Data Wrangling and visualization 

### 1. Display the first 10 rows of the dataset using `kable()` function
```{r}
kable((data[1:10, 1:5])) # first 10 rows of the dataset and the first 5 variables using kabble() function
```

### 2. Using `skim()` display the summary of variables
```{r}
data %>% 
  skim() # Pipe dataset into skim function
```

### 3. Use `fct_reorder` and `parse_number` functions to create a factor variable `family_income`
```{r}
# use fct_reorder
test_data <- data %>%
  mutate(family_income = fct_reorder(family_income, parse_number(family_income, na = c("Prefer not to answer")))) 



head(test_data)
rm(test_data)

# use arrange
arrange_data <- data %>%
  mutate(family_income = parse_number(family_income, na = c("Prefer not to answer"))) %>% #drops any non-numeric characters
  mutate(family_income=as_factor(family_income)) %>% #convert family_income to a factor
  arrange(family_income) 
```


### 4. What is the number of people who celebrate?
```{r}
data %>% 
  count(celebrate) %>% # Pipe dataset into count function and count restaurants.
  arrange(-n) %>%
  kable()
```

### 5. What are categories and insights for each main dish served and the method it is prepared?
```{r}
data %>%
  filter(!is.na(main_dish) & !is.na(main_prep)) %>%
  group_by(main_dish, main_prep) %>%
  count(main_dish, sort=TRUE) %>%
  kable()

```

### 6. Create 3 different data viz showing insights for main dish served and the method.

```{r}
data %>% 
  filter(!is.na(main_dish) & !is.na(main_prep)) %>%
  mutate(main_dish = fct_collapse(main_dish, Other = c("I don't know", "Other (please specify)", "Turducken"))) %>%
  mutate(main_prep = fct_collapse(main_prep, Other = c("I don't know", "Other (please specify)"))) %>%
  ggplot(aes(main_prep,color=main_prep,fill = main_prep))+ 
  geom_bar()+ # add Bar charts using geom_bar
  facet_wrap(vars(main_dish), ncol = 3, scales = "free_x")+ 
  labs(x = "Method") +
  coord_flip()+
  theme(legend.position = "none")
```


```{r}
data %>%
  filter(!is.na(main_dish) & !is.na(main_prep)) %>%
  mutate(main_dish = fct_collapse(main_dish, Other = c("I don't know", "Other (please specify)", "Turducken"))) %>%
  mutate(main_prep = fct_collapse(main_prep, Other = c("I don't know", "Other (please specify)"))) %>%
  ggplot(aes(main_dish,color=main_dish,fill = main_dish))+
  geom_bar()+
  facet_wrap(vars(main_prep), ncol = 2, scales = "free_x") +
  labs(x = "Dish served") +
  coord_flip()
```


```{r}
# Use fct_lump to lump together all the small groups 
data %>%
  filter(!is.na(main_dish) & !is.na(main_prep)) %>%
  mutate(main_dish = fct_lump(main_dish)) %>%
  mutate(main_prep = fct_lump(main_prep)) %>%
  ggplot(aes(main_dish,color=main_dish,fill = main_dish))+
  geom_bar()+
  facet_wrap(vars(main_prep), ncol = 3)+
  labs(x = "Dish served") +
  theme(legend.position = "none")
```


### 7. How many use cranberry sauce? How many use gravy?
```{r}
data %>%
  filter(!is.na(main_dish)) %>%
  mutate(cranberry = fct_collapse(cranberry,
                                  Yes = c("Homemade","Canned", "Other (please specify)"),
                                  No = c("None"))) %>%
  count(cranberry,gravy, sort = TRUE)
```


```{r}
data %>%
  filter(!is.na(main_dish)) %>%
  mutate(cranberry = fct_collapse(cranberry,
                                  Yes = c("Homemade","Canned", "Other (please specify)"),
                                  No = c("None"))) %>%
  ggplot(aes(cranberry,color=cranberry,fill = cranberry))+
  geom_bar()+
  theme(legend.position = "none")
```

```{r}
data %>%
  filter(!is.na(main_dish)) %>%
  mutate(gravy = gravy %>% fct_infreq() %>% fct_rev()) %>%
  ggplot(aes(gravy)) +
  geom_bar()
```

### 8. What is the distribution of those who celebrate across income ranges

```{r}
# Use fct_lump to lump together all the small groups 
data %>%
  mutate(family_income = fct_lump(family_income)) %>%
  ggplot(aes(family_income,color=family_income,fill = family_income))+
  geom_bar()+
  facet_wrap(vars(celebrate), ncol = 8)+
  labs(x = "Income") +
  coord_flip() +
  theme(legend.position = "none")

data %>%
  filter((celebrate=="Yes") & !is.na(family_income)) %>%
  #mutate(family_income = fct_lump(family_income)) %>%
  ggplot(aes(family_income,color=family_income,fill = family_income))+
  geom_bar()+
  labs(x = "Income") +
  coord_flip() +
  theme(legend.position = "none")
```

### 10. Use the following code to create a new data 

```{r}
new_data <- data %>%
  select(id, starts_with("side"),
         starts_with("pie"),
         starts_with("dessert")) %>%
  select(-side15, -pie13, -dessert12) %>%
  gather(type, value, -id) %>%
  filter(!is.na(value),
         !value %in% c("None", "Other (please specify)")) %>%
  mutate(type = str_remove(type, "\\d+"))
```

### Intall package `widyr` and use `pairwise_cor()` function
```{r}
install.packages("widyr")
```




```{r}
data %>%
  filter(!is.na(main_dish)) %>%
  mutate(main_dish = fct_collapse(main_dish,
                                  Other = c("I don't know", "Other (please specify)")
  )) %>%
  count(main_dish,main_dish_other, sort = TRUE)
```



