---
title: "Assessment 1.1"
author: "Christopher Schmidt S4532362"
date: "15/08/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RMarkdown editing

**R Markdown** contains a combination of markdown (*content with simple text formatting*) and R code chunks.

R Markdown documents allow you to weave together narrative text and code to produce elegantly formatted output in dozens of formats, such as HTML, PDF, and MS Word documents. 

For more details on how to use R Markdown see [Markdown Basics](https://rmarkdown.rstudio.com/lesson-8.html)

This document was created with **R Markdown**.

```{r}
## install packages once via console using - install.packages("tidyverse")
library(tidyverse) # load tidyverse library
library(skimr)     # load skimr library
library(janitor)   # load janitor library
```

### Insert an R chunk and create a dataset variable called `fastfood`
```{r}
fastfood <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv')
```

### Display the **first 10 rows** and **first 5 variables** of the `fastfood` dataset
```{r}
# fastfood%>%head(10) # Display the first 10 rows of the dataset using head() function.
knitr::kable(head(fastfood[, 1:5])) # first 10 rows of the dataset and the first 5 variables using kabble() function
```

### Display observations that have more than **1000 calories**
```{r}
fastfood%>%             # Pipe dataset into filter function 
  filter(calories>1000) # Filter by calories greater than 1000
```

### Save observations with more than **40 in total_fat** and more than **80 in total_carb** in a new variable called `dont_eat_this`
```{r}
dont_eat_this <- fastfood%>%           # Create new variable called `dont_eat_this` and assign filtered dataset to it 
  filter(total_fat>40 & total_carb>80) # filter by total_fat greater than 40 & total_carb greatre than 80
```

### Using the initial dataset variable, display observations from restaurant **Mcdonalds**
```{r}
fastfood%>%                       # Pipe dataset into filter function
  filter(restaurant=="Mcdonalds") # Filter by restaurant named Mcdonalds
```




