---
title: "Assessment 1.1"
author: "Christopher Schmidt S4532362"
date: "15/08/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RMarkdown editing

**R Markdown** contains a combination of markdown (*content with simple text formatting*) and R code chunks.

R Markdown documents allow you to weave together narrative text and code to produce elegantly formatted output in dozens of formats, such as HTML, PDF, and MS Word documents. 

For more details on how to use R Markdown see [Markdown Basics](https://rmarkdown.rstudio.com/lesson-8.html)

This document was created with **R Markdown**.

```{r}
library(tidyverse)
library(janitor)
library(skimr)

fastfood <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv')

fastfood <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv')

fastfood%>%head(10) #Display the first 10 rows of the dataset.

knitr::kable(head(fastfood[, 1:5]))

fastfood%>%
  filter(calories>1130) #Display the observations that has more than 1000 calories
```


